---
title: "Lab3"
author: "Your Name"
date: "2024-09-19"
output: "github_document"
---

# Load Packages
```{r}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)
library(ggsci)
library(see)

```

# Load Datasets
```{r}

lab3data <- read.csv("/Users/northlight/Documents/GitHub/Lab3/lab3data.csv")

```


# Missing Data
```{r}

#Listwise deletion
list_lab3data<- drop_na(lab3data)

#pair wise deletion will only be used when you run the actual analysis using "na.rm = TRUE"

```

# Recode Variables: Changing Numeric Values to Characters
```{r}

lab3data %>% count(Gender)

#Solution 1
lab3data<- lab3data%>%
  mutate_at(c('Gender'),funs(str_replace(., "1", "Male")))

lab3data<- lab3data%>%
  mutate_at(c('Gender'),funs(str_replace(., "2", "Female")))

lab3data %>% count(Gender)


#Solution 2

#lab3data$Gender<-recode(lab3data$Gender, '1' = 'X', '2' = 'Y')


#Now recode 2 into Female




```

# Recode Variables: Changing Numeric Values to Different Numeric Values
```{r}

lab3data %>% count(RSE2)

lab3data$RSE2_R <- 5 - lab3data$RSE2

#check to see if the recoding was successful
lab3data %>% count(RSE2_R)

#Now do the same for the other items (5,6,8,9)

lab3data %>% count(RSE5)

lab3data$RSE5_R <- 5 - lab3data$RSE5

lab3data %>% count(RSE6)

lab3data$RSE6_R <- 5 - lab3data$RSE6

lab3data %>% count(RSE8)

lab3data$RSE8_R <- 5 - lab3data$RSE8

lab3data %>% count(RSE9)

lab3data$RSE9_R <- 5 - lab3data$RSE9

```


# Create Composites
```{r}

#Depending on the scale, sometimes you need to average across the items, other times you might need to sum 


#Fill in the items
lab3data <- lab3data %>%
  mutate(RSE = rowMeans(cbind(RSE1, RSE2_R, RSE3, RSE4, RSE5_R, RSE6_R, RSE7, RSE8_R, RSE9_R, RSE10)))


#Create another composite for SWL, but this time use rowSums

lab3data <- lab3data %>%
  mutate(SWL = rowSums(cbind(SWL1, SWL2, SWL3, SWL4, SWL5)))


```


# Get Summary Descriptive Statistics
```{r}

lab3data %>% 
  group_by(Gender) %>%
  dplyr::summarize(mean_RSE = mean(RSE), 
      mean_SWL = mean(SWL),
      std_dev_RSE = sd(RSE),
      std_dev_SWL = sd(SWL),
      corr_RSE_SWL  = cor(RSE, SWL))




```

# Q1: Descrptively, which group has a higher self-esteem? What about life satisfaction? Is the relationship between self-esteem and life satisfaction stronger for males or for females? 

Males appear to have higher self-esteem. 
Females appear to have higher life satisfaction. 
The relationship between self esteem and life satisfaction appears to be stronger for Males. 

# Visualization of Data Distribution
```{r}

lab3data<- lab3data %>%
  drop_na(Gender)


ggplot(lab3data, aes(x = Gender, y = RSE)) + geom_violin(aes(fill=Gender)) + scale_fill_simpsons() + theme_modern()

#Create a plot for SWL by Gender with a different color palette

ggplot(lab3data, aes(x = Gender, y = SWL)) + geom_violin(aes(fill=Gender)) + scale_fill_tron() + theme_modern()

?scale_fill_brewer

```

# Q2: What you can tell from looking at a violin plot? What you cannot tell? What inferences you can make based on the two violin plots? 

A violin plot itself can show us: 
- Distribution of scores; 
- Mode (ie in this case, most responders, both male and female in SWE fall at 4/5 for life satisfaction) 
- Total possible Range of scores (ie 0-25 here)

A violin plot itself cannot show us: 
- Mean
- Median
- Standard Deviation
- Standard Error 
- Total number of scores from each subject 
- Total number of surveyed subjects
- correlation (separate plots)

We can infer from this violin plot: 
We see that Male and Female distribution appears to be quite visually similar in terms of life satisfaction, however, there are a couple subtle visual cues; more female than male subjects fall at 10/25 and 15/25for total life satisfaction score, whereas more males appear to land at 20/25. Also, it appears that slightly more male respondents fall at the 5/25 total score than females. Could it be that Males are more likely to fall in the extremes of satisfaction, whereas Females are more likely to be middle of the road? 

In terms of self esteem, again they do appear to have visually similar distributions, with some slight differences; visual cues for difference show a more triangular distribution for Males, whereas Females show a more spherical distribution. There appear to be two modes for Male subjects at 3 and at 4 (or slightly below 4) whereas the smoother distribution for Female subjects does not indicate two modes; it appears that the Female mode lands around 3, but no strong peaks or valleys.
